---
title: "Flux Data Summary"
author: "Trevor Williams"
date: "September 4, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, conflicts = FALSE)
```

## N2O Flux Data

This is a preliminary report on **N2O Flux Data** from the Willamette project. Many sources were referenced for the various tests employed including <https://rcompanion.org/handbook/index.html>. 

This document will serve as a template for initial reporting of data from all fields and plots, including CO2 flux and water anlyses.

```{r, echo=FALSE}

# package and library load
library(tidyverse)
#import flux data
flux_data <- read_csv("C:/Users/twilli2/Dropbox/Lab data/S Willamette GWMA Dropbox/Flux Data/Slopes and Flux CSV.csv",
col_types = cols(chamber = col_character(),
  compound = col_factor(levels = c("co2",
"n2o")), date = col_datetime(format = "%m/%d/%Y"),
field = col_factor(levels = c("1","2","3","4")), plot = col_factor(levels = c("0",
"25", "50", "75", "100", "C",
"P"))))
.libPaths()
#n2o flux data summary

flux_data_n2o <- filter(flux_data, compound =="n2o")

library(tidyverse)
spread(flux_data, plot, flux) %>% 
filter(compound == "n2o") %>% 
summary()

```
```{r}
#Conv VS EEF
flux_data_n2o_cp <- filter(flux_data_n2o, plot == "C"| plot == "P") 

ggplot(data = flux_data_n2o_cp, mapping = aes(x = flux))+
  geom_histogram(binwidth = 0.5)

ggplot(data = flux_data_n2o_cp, mapping = aes(x = flux, color = plot))+
  geom_freqpoly(binwidth = 0.5)

shapiro.test(flux_data_n2o_cp$flux)#null hyp = data from norm distribution. Rejected
```
Null hypothesis for Shapiro-Wilk test, that data are from a normal distribution is rejected.
First lets focus on the the Conventional VS Precicion Agriculture or Enhanced Efficiency Fertilizer

##Conv. VS P.A.

```{r}
ggplot(flux_data_n2o_cp, aes(x = plot, y = flux, color = plot)) +
  geom_boxplot(outlier.alpha = 0.5, notch = T, varwidth = T, 
               fill = "white", show.legend = T)+
  labs(x = "Plot", y = "Flux (ng N/cm2/hr)")+
  facet_grid(~field)

ggplot(flux_data_n2o_cp, aes(x = plot, y = flux, color = plot)) +
  geom_boxplot(outlier.alpha = 0.1, notch = T, varwidth = T, 
               fill = "white", show.legend = T)+
  labs(x = "Plot", y = "Flux (ng N/cm2/hr)")+
  coord_cartesian(ylim = c(-4, 8))+
  facet_grid(~field)
```

First boxplot shows full range of data with outliers.

Second boxplot shows the central tendency of the data clearly, with 50% of the data indicated by the box, the median clearly visible as a line in the middle, and the IQ range +/- 1.5
indicated by the whiskers.

***

####T-tests are for normally distrubted data around the mean.
####Welch Two Sample t-Test for unequal variance

```{r}

t.test(flux~plot, data = flux_data_n2o_cp)


```
###Field 1
```{r}
#What they look like on each field individually
flux_data_n2o_cp_F1 <- filter(flux_data_n2o, plot == "C"| plot == "P", field == "1")
t.test(flux~plot, data = flux_data_n2o_cp_F1, var.equal = T)
```
###Field 2
```{r}
flux_data_n2o_cp_F2 <- filter(flux_data_n2o, plot == "C"| plot == "P", field == "2")
t.test(flux~plot, data = flux_data_n2o_cp_F2, var.equal = T)

```
###Field 3
```{r}
flux_data_n2o_cp_F3 <- filter(flux_data_n2o, plot == "C"| plot == "P", field == "3")
t.test(flux~plot, data = flux_data_n2o_cp_F3)

```
###Field 4
```{r}
flux_data_n2o_cp_F4 <- filter(flux_data_n2o, plot == "C"| plot == "P", field == "4")
t.test(flux~plot, data = flux_data_n2o_cp_F4)
#no significant difference between means, with 95% CI falling below means in both groups in all field
```
####No significant difference between means, with 95% CI falling below means in both groups in all field

|NONPARAMETRIC TEST	| PARAMETRIC ALTERNATIVE|
|------:|:-----|
|1-sample sign test	| One-sample Z-test, One sample t-test|
|1-sample Wilcoxon Signed Rank test	| One sample Z-test, One sample t-test|
|Friedman test	| Two-way ANOVA|
|Kruskal-Wallis test |	One-way ANOVA|
|Mann-Whitney test	| Independent samples t-test
|Mood’s Median test	| One-way ANOVA
|Spearman Rank Correlation	| Correlation Coefficient

###Wilcox tests
####Results of Wilcoxon rank sum test on all fields and each field individually
All fields
```{r}
wilcox.test(flux~plot, data = flux_data_n2o_cp) 
```
Significant, p-value < 0.05

Field 1
```{r}
wilcox.test(flux~plot, data = flux_data_n2o_cp_F1) #approaching significance
```
Field 2
```{r}
wilcox.test(flux~plot, data = flux_data_n2o_cp_F2)
```
Field 3
```{r}
wilcox.test(flux~plot, data = flux_data_n2o_cp_F3)
```
Significant, p-value < 0.05

Field 4
```{r}
wilcox.test(flux~plot, data = flux_data_n2o_cp_F4)
```

###Median Tests

```{r}
if(!require(psych)){install.packages("psych")}
if(!require(FSA)){install.packages("FSA")}
if(!require(lattice)){install.packages("lattice")}
if(!require(multcompView)){install.packages("multcompView")}
if(!require(rcompanion)){install.packages("rcompanion")}
if(!require(Rcpp)){install.packages("Rcpp")}

library(lattice)
library(FSA)
library(RVAideMemoire)
#Summarise CP Median Data

Summarize(flux ~ plot,
          data = flux_data_n2o_cp,
          digits = 3)
```

### Median Test for all fields

```{r}
mood.medtest(flux ~ plot,
             data  = flux_data_n2o_cp,
             exact = FALSE)
flux_data_n2o_cp$plot = factor(flux_data_n2o_cp$plot, 
                      levels=c("C","P"))

PT = pairwiseMedianMatrix(flux ~ plot,
                        data   = flux_data_n2o_cp,
                        exact  = NULL,
                        method = "fdr")

library(multcompView)
PT
final <- multcompLetters(PT$Adjusted,
                compare="<",
                threshold=0.05,
                Letters=letters)
final
```
The median of the C plots is significantly higher than the median of the P plots, overall.

###Field 1

```{r}
mood.medtest(flux ~ plot,
             data  = flux_data_n2o_cp_F1,
             exact = FALSE)
flux_data_n2o_cp_F1$plot = factor(flux_data_n2o_cp_F1$plot, 
                      levels=c("C","P"))

PT = pairwiseMedianMatrix(flux ~ plot,
                        data   = flux_data_n2o_cp_F1,
                        exact  = NULL,
                        method = "fdr")


PT
final <- multcompLetters(PT$Adjusted,
                compare="<",
                threshold=0.05,
                Letters=letters)
final

Summarize(flux ~ plot,
          data = flux_data_n2o_cp_F1,
          digits = 3)
```

In field 1, the median of the C plot is significantly higher than the median of the P plot

###Field 2
```{r}

mood.medtest(flux ~ plot,
             data  = flux_data_n2o_cp_F2,
             exact = FALSE)
flux_data_n2o_cp_F2$plot = factor(flux_data_n2o_cp_F2$plot, 
                      levels=c("C","P"))

PT = pairwiseMedianMatrix(flux ~ plot,
                        data   = flux_data_n2o_cp_F2,
                        exact  = NULL,
                        method = "fdr")


PT
final <- multcompLetters(PT$Adjusted,
                compare="<",
                threshold=0.05,
                Letters=letters)
final
Summarize(flux ~ plot,
          data = flux_data_n2o_cp_F2,
          digits = 3)
```

No significant difference in plot medians of Field 2

###Field 3

```{r}
mood.medtest(flux ~ plot,
             data  = flux_data_n2o_cp_F3,
             exact = FALSE)
flux_data_n2o_cp_F3$plot = factor(flux_data_n2o_cp_F3$plot, 
                      levels=c("C","P"))

PT = pairwiseMedianMatrix(flux ~ plot,
                        data   = flux_data_n2o_cp_F3,
                        exact  = NULL,
                        method = "fdr")


PT
final <- multcompLetters(PT$Adjusted,
                compare="<",
                threshold=0.05,
                Letters=letters)
final
Summarize(flux ~ plot,
          data = flux_data_n2o_cp_F2,
          digits = 3)
```

Median of C is significantly higher than median of P in field 3


###Field 4

```{r}
mood.medtest(flux ~ plot,
             data  = flux_data_n2o_cp_F4,
             exact = FALSE)
flux_data_n2o_cp_F4$plot = factor(flux_data_n2o_cp_F4$plot, 
                      levels=c("C","P"))

PT = pairwiseMedianMatrix(flux ~ plot,
                        data   = flux_data_n2o_cp_F4,
                        exact  = NULL,
                        method = "fdr")


PT
final <- multcompLetters(PT$Adjusted,
                compare="<",
                threshold=0.05,
                Letters=letters)
final
Summarize(flux ~ plot,
          data = flux_data_n2o_cp_F4,
          digits = 3)
```

No significant differences in medians of plots in Field 4

Both with the Wilcox Test and the Mood test, fields 1 and 3 have higher central tendencies in the C plots versus the P plots.

##Gradient Plots

```{r}
flux_data_n2o_g <- filter(flux_data_n2o, plot != "C" & plot !="P") # Gradient Data
ggplot(data = flux_data_n2o_g, mapping = aes(x = flux))+
  geom_histogram(binwidth = 0.5)

ggplot(data = flux_data_n2o_g, mapping = aes(x = flux, color = plot))+
  geom_freqpoly(binwidth = 0.5)

shapiro.test(flux_data_n2o_g$flux)#null hyp = data from norm distribution. Rejecte
```

####Null hypothesis that data are from normal distribution is rejected.

```{r}
Summarize(flux_data_n2o_g$flux)

ggplot(flux_data_n2o_g, aes(x = plot, y = flux, color = plot)) +
  geom_boxplot(outlier.alpha = 0.1, notch = T, varwidth = T, 
               fill = "white", show.legend = T)+
  labs(x = "Plot", y = "Flux (ng N/cm^2/hr)")+
  facet_grid(~field)

ggplot(flux_data_n2o_g, aes(x = plot, y = flux, color = plot)) +
  geom_boxplot(outlier.alpha = 0.1, notch = T, varwidth = T, 
               fill = "white", show.legend = T)+
  labs(x = "Plot", y = "Flux (ng N/cm^2/hr)")+
  coord_cartesian(ylim = c(-4, 10))+
  facet_grid(~field)

```

```{r}
if(!require(psych)){install.packages("psych")}
if(!require(FSA)){install.packages("FSA")}
if(!require(lattice)){install.packages("lattice")}
if(!require(multcompView)){install.packages("multcompView")}
if(!require(rcompanion)){install.packages("rcompanion")}
if(!require(Rcpp)){install.packages("Rcpp")}
library(lattice)
library(FSA)
library(coin)
g_F1 <- filter(flux_data_n2o_g, field == "1")
g_F2 <- filter(flux_data_n2o_g, field == "2")
g_F3 <- filter(flux_data_n2o_g, field == "3")
g_F4 <- filter(flux_data_n2o_g, field == "4")
g_F1
g_F2
g_F3
g_F4

```

###Median Tests

```{r}
median_test(flux~plot, data = flux_data_n2o_g)
```


###Field 1
```{r}
spread(g_F1, plot, flux) %>% 
summary()
```

```{r}
median_test(flux~plot, data = g_F1) #p-value < 0.05, “There was a significant difference in the median values among groups.”
#Field 2
median_test(flux~plot, data = g_F2)
#Field 3
median_test(flux~plot, data = g_F3)
#Field 4
median_test(flux~plot, data = g_F4)

```

In all fields, p-value < 0.05, “There was a significant difference in the median values among groups.”

###Field 1

```{r}
library(RVAideMemoire)
#field 1
mood.medtest(flux ~ plot,
             data  = g_F1,
             exact = FALSE)
g_F1$plot = factor(g_F1$plot, 
                      levels=c("0", "25", "50", "75", "100"))

PT = pairwiseMedianMatrix(flux ~ plot,
                        data   = g_F1,
                        exact  = NULL,
                        method = "fdr")

library(multcompView)
PT
multcompLetters(PT$Adjusted,
                compare="<",
                threshold=0.05,
                Letters=letters)
```

```{r}
ggplot(g_F1, aes(x = plot, y = flux, color = plot)) +
  geom_boxplot(outlier.alpha = 0.1, notch = T, varwidth = T, 
               fill = "white", show.legend = T)+
  labs(x = "Plot", y = "Flux (ng N/cm^2/hr)")+
  coord_cartesian(ylim = c(-4, 10))
```

###Kruskal-Wallis non-parametric alternative to one-way Anova with more than two groups

```{r}
library(ggpubr)
compare_means(flux ~ plot, data = g_F1)
my_comparisons <- list( c("0", "75"), c("0", "100"), c("25", "75"), c("25","100") )
ggboxplot(g_F1, x = "plot", y = "flux",
          color = "plot")+ 
  labs(x = "Plot", y = "Flux (ng N/cm^2/hr)")+
  coord_cartesian(ylim = c(-4, 10))+
  stat_compare_means(comparisons = my_comparisons, label.y = c(2, 4.25, 6.5, 10), bracket.size = 0.1, tip.length = 0.001)+
  stat_compare_means(label.y = 9.5
                     )
flux_sum <- g_F1 %>% 
  group_by(plot) %>% 
  summarise(meanf = mean(flux, na.rm = T),
            sd_f = sd(flux, na.rm = T),
            n_f = n(),
            SE_f = sd_f/sqrt(n_f))

ggplot(flux_sum, aes(plot, meanf, color = plot))+
  geom_col(fill = "white")+
  geom_errorbar(aes(ymin = meanf - SE_f, ymax = meanf + SE_f), width = 0.2)+
  labs(x = "Plot", y = "Mean flux (ng N/cm^2/hr)")+
  #stat_compare_means(data = g_F1, aes(x = plot, y = flux), comparisons = my_comparisons, label.y = c(2, 4.25, 6.5, 10), bracket.size = 0.1, tip.length = 0.001)+
  stat_compare_means(data = g_F1, aes(x = plot, y = flux), label.y = 3.5)

```

###Field 2

```{r}
#field 2
mood.medtest(flux ~ plot,
             data  = g_F2,
             exact = FALSE)
g_F2$plot = factor(g_F2$plot, 
                      levels=c("0", "25", "50", "75", "100"))
library(rcompanion)
PT = pairwiseMedianMatrix(flux ~ plot,
                        data   = g_F2,
                        exact  = NULL,
                        method = "fdr")

library(multcompView)
multcompLetters(PT$Adjusted,
                compare="<",
                threshold=0.05,
                Letters=letters)
#medians are significantly different
#Kruskal-Wallis for means
compare_means(flux ~ plot, data = g_F2)
my_comparisons <- list( c("0", "50"), c("0", "75"), c("0", "100"), c("25","50"), c("25","75"), c("25","100") )
ggboxplot(g_F2, x = "plot", y = "flux",
          color = "plot") +
          #add = "jitter")+ 
  labs(x = "Plot", y = "Flux (ng N/cm^2/hr)")+
  #coord_cartesian(ylim = c(-4, 10))+
  stat_compare_means(comparison = my_comparisons, label = "p.signif", step.increase = 0.5)+
  stat_compare_means(label.y = 10)
flux_sum <- g_F2 %>% 
  group_by(plot) %>% 
  summarise(meanf = mean(flux, na.rm = T),
            sd_f = sd(flux, na.rm = T),
            n_f = n(),
            SE_f = sd_f/sqrt(n_f))

ggplot(flux_sum, aes(plot, meanf, color = plot))+
  geom_col(fill = "white")+
  geom_errorbar(aes(ymin = meanf - SE_f, ymax = meanf + SE_f), width = 0.2)+
  labs(x = "Plot", y = "Mean flux (ng N/cm^2/hr)")+
  #stat_compare_means(data = g_F1, aes(x = plot, y = flux), comparisons = my_comparisons, label.y = c(2, 4.25, 6.5, 10), bracket.size = 0.1, tip.length = 0.001)+
  stat_compare_means(data = g_F2, aes(x = plot, y = flux), label.y = 3.5)
```

###Field 3

```{r}
#field 3
mood.medtest(flux ~ plot,
             data  = g_F3,
             exact = FALSE)
g_F3$plot = factor(g_F3$plot, 
                      levels=c("0", "25", "50", "75", "100"))
library(rcompanion)
PT = pairwiseMedianMatrix(flux ~ plot,
                        data   = g_F3,
                        exact  = NULL,
                        method = "fdr")

library(multcompView)
multcompLetters(PT$Adjusted,
                compare="<",
                threshold=0.05,
                Letters=letters)

compare_means(flux ~ plot, data = g_F3)
my_comparisons <- list( c("0", "50"), c("0", "75"), c("0", "100"), c("25","75"), c("25","100"),c("50","75"),c("50","100") )
ggboxplot(g_F3, x = "plot", y = "flux",
          color = "plot") +
          #add = "jitter")+ 
  labs(x = "Plot", y = "Flux (ng N/cm^2/hr)")+
  #coord_cartesian(ylim = c(-4, 10))+
  stat_compare_means(comparison = my_comparisons, label = "p.signif", step.increase = 0.5)+
  stat_compare_means(label.y = 10)

ggboxplot(g_F3, x = "plot", y = "flux",
          color = "plot") +
          #add = "jitter")+ 
  labs(x = "Plot", y = "Flux (ng N/cm^2/hr)")+
  coord_cartesian(ylim = c(-4, 10))+
  stat_compare_means(comparison = my_comparisons, label = "p.signif", step.increase = 0.5)+
  stat_compare_means(label.y = 10)

flux_sum <- g_F3 %>% 
  group_by(plot) %>% 
  summarise(meanf = mean(flux, na.rm = T),
            sd_f = sd(flux, na.rm = T),
            n_f = n(),
            SE_f = sd_f/sqrt(n_f))

ggplot(flux_sum, aes(plot, meanf, color = plot))+
  geom_col(fill = "white")+
  geom_errorbar(aes(ymin = meanf - SE_f, ymax = meanf + SE_f), width = 0.2)+
  labs(x = "Plot", y = "Mean flux (ng N/cm^2/hr)")+
  #stat_compare_means(data = g_F1, aes(x = plot, y = flux), comparisons = my_comparisons, label.y = c(2, 4.25, 6.5, 10), bracket.size = 0.1, tip.length = 0.001)+
  stat_compare_means(data = g_F3, aes(x = plot, y = flux), label.y = 3.5)
```

###Field 4

```{r}
#field 4
mood.medtest(flux ~ plot,
             data  = g_F4,
             exact = FALSE)
g_F4$plot = factor(g_F4$plot, 
                      levels=c("0", "25", "50", "75", "100"))
library(rcompanion)
PT = pairwiseMedianMatrix(flux ~ plot,
                        data   = g_F4,
                        exact  = NULL,
                        method = "fdr")

library(multcompView)
multcompLetters(PT$Adjusted,
                compare="<",
                threshold=0.05,
                Letters=letters)

compare_means(flux ~ plot, data = g_F4)
my_comparisons <- list( c("0", "50"), c("0", "75"), c("0", "100"), c("25","75"), c("25","100"), c("50","75"), c("50", "100"))
ggboxplot(g_F3, x = "plot", y = "flux",
          color = "plot") +
          #add = "jitter")+ 
  labs(x = "Plot", y = "Flux (ng N/cm^2/hr)")+
  #coord_cartesian(ylim = c(-4, 10))+
  stat_compare_means(comparison = my_comparisons, label = "p.signif", step.increase = 0.5)+
  stat_compare_means(label.y = 10)

ggboxplot(g_F4, x = "plot", y = "flux",
          color = "plot") +
          #add = "jitter")+ 
  labs(x = "Plot", y = "Flux (ng N/cm^2/hr)")+
  coord_cartesian(ylim = c(-4, 10))+
  stat_compare_means(comparison = my_comparisons, label = "p.signif", step.increase = 0.5)+
  stat_compare_means(label.y = 10)

flux_sum <- g_F4 %>% 
  group_by(plot) %>% 
  summarise(meanf = mean(flux, na.rm = T),
            sd_f = sd(flux, na.rm = T),
            n_f = n(),
            SE_f = sd_f/sqrt(n_f))

ggplot(flux_sum, aes(plot, meanf, color = plot))+
  geom_col(fill = "white")+
  geom_errorbar(aes(ymin = meanf - SE_f, ymax = meanf + SE_f), width = 0.2)+
  labs(x = "Plot", y = "Mean flux (ng N/cm^2/hr)")+
  #stat_compare_means(data = g_F1, aes(x = plot, y = flux), comparisons = my_comparisons, label.y = c(2, 4.25, 6.5, 10), bracket.size = 0.1, tip.length = 0.001)+
  stat_compare_means(data = g_F4, aes(x = plot, y = flux), label.y = 3.5)
```

We have a pattern of N2O median and mean flux significantly higher in the 75 and 100 percent plots as compared with the 0, 25, and 50 percent plots.

Next we will look at differences between field soil characteristics, and any influence of soil moisture and temperature on flux dynamics.

```{r}
ggplot(data = flux_data_n2o_g, aes(date,flux, color = plot, shape = field))+
  geom_point()+
  geom_line(aes(date,soiltemp),color = "black")
```

